### 2021-01-07 3_r_scripts/1_markdown_document.Rmd
Setting up project.

---
title: "Can we distinguish between the speed and scope of acute glucocorticoid responses?"
author: "Conor Taff, Laura Schoenle, and Maren Vitousek"
date: "Last Updated: `r format(Sys.time(), '%B %d %Y at %H:%M')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(ggplot2)
library(here)
library(sjPlot)
```

```{r}
sessionInfo()
```